Title

### 2022-05-03 thesis/phases/phases.tex
phases

The energy efficiency of HPC can be achieved through dynamic workload planning. Dynamic scaling of frequency and voltage DVFS is the main energy efficiency technique to save power. The scaling algorithms can operate dynamically depending on the CPU workload. However, the required IT resources and workload vary throughout the applications' execution. The division of this workload directly impacts the resulting performance and the complexity of dynamic scaling algorithms. This work analyses the impact on the energy efficiency of different dynamic workload partitioning techniques, based on time division of allocated resources and DVFS. We evaluate the trade-off between the best time intervals partition and the overhead impact of dynamic scheduling techniques. We propose the optimal partition of application executions in time intervals providing the most energy efficient execution parameters, workload, frequency and number of cores, for each time interval.

\section{Introduction to models on time} \label{sec:introduction_to_models_on_time}
Talk about the importance of energy efficiency.

>>>>

Dynamic frequency scaling and dynamic power management are the leading software techniques for energy saving in modern computers. These techniques work through dynamic workload planning, and switching from one power state to another is crucial for them. This work analyses the impact on the energy efficiency of different dynamic workload partitioning techniques based on the time-division and allocated resources. We propose a methodology combining measurement data with a simulation to provide insights into choosing the best time intervals. Furthermore, we evaluate the trade-off between the best time intervals partition and the overhead impact of dynamic scheduling techniques identifying that generally exists a limit to the number of divisions that the algorithm can benefit from, giving a lower bound to the minimal time interval.

\section{Introduction to models on time} \label{sec:introduction_to_models_on_time}
Energy efficiency in computing has always been a concern, especially in high-performance computing (HPC), where a small percentage in energy savings can significantly reduce maintenance costs or environmental impact due to its scale of consumption. For example, the leading Petaflop supercomputers consume are in a range of 1–18 MW of electrical power, with 1.5 MW on average, which we can estimate to be into a millions dollars range of electricity bills annually \cite{Group2012HandbookSahni}. Also,  the International Energy Agency  \cite{iea_2021} estimated that global data center electricity use in 2020 was around 200-250 TWh, or around 1\% of the global electricity demand, which can generate as much pollution as a nation like Argentina \cite{Mathew2012Energy-awareNetworks}.

Given the importance of this topic, modern processors of desktop, mobile phones, and HPC  have internal power management mechanisms, and that is because the processor is one of the main components that drain energy and can reach 50\% of the total system consumption \cite{Fan2007, Barroso2007TheComputing, Malladi2012TowardsDRAM}. Within the mechanisms implemented by the processors, the main controllable via software are Dynamic Frequency and Voltage Scaling (DVFS) and Dynamic Power Management (DPM) ~\cite{Rotem2012Power-managementBridge, Brown2005, Hackenberg2015}.

DPM encompasses a set of techniques that take advantage of different power levels on the system, such as active, idle, and disabled~\cite{CARDOSO201717, Shuja2012Energy-efficientCenters, Benini2000AManagement}. The basic idea is to shut down devices and wake them up when necessary. However, this is not trivial, that are several variables that need to take into account as the switching cost, in terms of time and energy, for how long it will remain in that state,  the power consumption on every state, and a lot more.

The DPM technique is quite interesting on systems where static power is very high or where the system is idle for a long time. In these cases, some works report savings of up to 70\% in energy~\cite{Shuja2012Energy-efficientCenters, Benini2000AManagement}.

DVFS allows real-time control of the frequency and voltage depending on the needs. This method is motivated by the fact that we can approximate the relationship of the frequency with the power as a cubic equation and frequency and performance as a linear equation \cite{Dayarathna2016DataSurvey, Group2012HandbookSahni}, implying that a reduction in frequency has a cubic impact on power and linear on performance. Moreover, reducing the frequency leads to a longer execution time, increasing energy consumption. As a result, determining the best voltage and frequency to employ in all conditions is a challenging task.
